---
title: "Operations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Operations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(bignum)
```

## Recycling rules

The two inputs have compatible sizes if:

* Both input vectors have the same length, or
* One input is a scalar (length 1) which is recycled.


## Comparison Operations

`biginteger()` and `bigfloat()` vectors support the standard comparison operations. The base R documentation for these can be found via `help("Comparison")`.

```{r, eval=FALSE}
x < y
x > y
x <= y
x >= y
x == y
x != y
```

These operations always return a logical vector.


## Arithmetic Operations

`biginteger()` and `bigfloat()` vectors support the standard arithmetic operations. The base R documentation for these can be found via `help("Arithmetic")`.

```{r, eval=FALSE}
+ x
- x
x + y
x - y
x * y
x / y
x ^ y
x %% y
x %/% y
```

These arithmetic operations are **type-stable**, which means the output type depends only on the input types (not the input values). A biginteger vector is returned when the result must be an integer (e.g., addition of two integers). Otherwise a bigfloat vector is returned.

The following table summarizes the return type for each combination, where "integer-like" refers to integer and biginteger vectors and "float-like" refers to double and bigfloat vectors.

| **Input 1**  | **Operator**      | **Input 2**  |     | **Result** |
|:-------------|:-----------------:|:-------------|:---:|:-----------|
| Integer-like | +, -, \*, \^, %%  | Integer-like | -\> | biginteger |
| Integer-like | +, -, \*, \^, %%  | Float-like   | -\> | bigfloat   |
| Float-like   | +, -, \*, \^, %%  | Integer-like | -\> | bigfloat   |
| Float-like   | +, -, \*, \^, %%  | Float-like   | -\> | bigfloat   |
| Any          | /                 | Any          | -\> | bigfloat   |
| Any          | %/%               | Any          | -\> | biginteger |


## Mathematical Operations

`biginteger()` and `bigfloat()` vectors support the following mathematical operations:

```{r, eval=FALSE}
# special values
is.finite(x)
is.infinite(x)
is.nan(x)

# summary
sum(x)
prod(x)
max(x)
min(x)
range(x)
mean(x)

# cumulative
cumsum(x)
cumprod(x)
cummax(x)
cummin(x)

# rounding
floor(x)
ceiling(x)
trunc(x)

# miscellaneous
abs(x)
sign(x)
sqrt(x)

# logarithms and exponentials
log(x)
log10(x)
log2(x)
log1p(x)
exp(x)
expm1(x)

# trigonometric
cos(x)
sin(x)
tan(x)
acos(x)
asin(x)
atan(x)
cospi(x)
sinpi(x)
tanpi(x)

# hyperbolic
cosh(x)
sinh(x)
tanh(x)
acosh(x)
asinh(x)
atanh(x)

# special functions
gamma(x)
lgamma(x)
```

